Diffusive Homeostasis in a Self-Organizing Recurrent Neural Network \\ \begin{large} Spatially dependent Interaction as a Determinant of Neural Activity and Plasticity \end{large}

### 2017-03-10 draft.tex
.

@@

\subsubsection{The Role of Firing Rate Heterogeneity---a Dichotomy between Heterogeneity and Control?}
Modeling homeostasis necessarily includes a form of error estimation, which requires setting some predefined target that the feedback control will attempt to achieve. Of course this raises the question of how to choose these homeostatic targets. As mentioned in the previous section, former versions of the SORN defined a single global target value. However, experimental studies have found that neuronal populations exhibit a strong heterogeneity within their individual activity. Many statistical analyses found particular evidence for a log-normal like distribution of firing rates \cite{Buzsaki_Fir_Rates_2014,Wohrer_Fir_Rates_2012}. Following these results, one approach of modeling this feature would be to randomly distribute individual targets according to the aforementioned statistics. A problem of this procedure is the absence of any particular deeper physiological motivation or causation. In the light of this conceptual problem, one could, due to a lack of better knowledge, claim that it is reasonable to define a single global target of activity.

@@

Combining these results in a set of equations, we found
\begin{align}
\dot{NO}(t) &= -\lambda NO + \frac{ \left( r_{IP} + \alpha \left( V_{t,\rm pop.}(t)-V_{t,0} \right) \right)  N_{\rm exc.}\frac{{Ca^{2+}_{\rm spike}}^3 \tau_{Ca^{2+}} \mathrm{ln} (2)}{3}}{L^2} + \sigma_{NO} \xi(t) \label{2d-sys-homeostasis1}\\
\dot{V}_{t,\rm pop.}(t) &= \frac{NO-NO_0}{NO_0 \tau_{V_t}}\;. \label{2d-sys-homeostasis2}
\end{align}
Through the coordinate transformations
\begin{align}
n &\equiv NO-NO_0 \label{simplif_coord_transf1}\\
\theta &\equiv V_{t,\rm pop.}(t)-V_{t,0} \label{simplif_coord_transf2}\\
\end{align}
we could simplify the dynamical system:
\begin{align}
\dot{n} &= -\lambda n + \frac{\gamma \alpha N_{\rm exc.}}{L^2}\theta + \sigma_{NO} \xi(t) \label{2d-sys-homeostasis-coord_transf1}\\
\dot{\theta} &= \frac{1}{NO_0 \tau_{V_t}} n \label{2d-sys-homeostasis-coord_transf2}
\end{align}
where we have additionally used
\begin{align}
\gamma &\equiv \frac{{Ca^{2+}_{\rm spike}}^3 \tau_{Ca^{2+}} \mathrm{ln} (2)}{3} \label{2d-sys-fixed-points1}\\
NO_0 &= \frac{nNOS_{\mathrm{ total},0}}{\lambda L^2} \label{2d-sys-fixed-points2}\\
nNOS_{\mathrm{ total},0} &= r_{IP} N_{\rm exc.} \cdot \gamma \; . \label{2d-sys-fixed-points3}
\end{align}
As stated in the beginning, we were interested in finding an analytic expression for the power spectrum of the mean excitatory threshold. We therefore calculated the Fourier transform of \eqref{2d-sys-homeostasis-coord_transf1} and \eqref{2d-sys-homeostasis-coord_transf2}, yielding
\begin{align}
i\omega f_n &= - \lambda f_n + \frac{\gamma \alpha N_{\rm exc.}}{L^2}f_\theta + \sigma_{NO} f_{\xi} \label{2d-sys-homeostasis-coord_transf_ft1}\\
i\omega f_\theta &= \frac{f_n}{NO_0\tau_{V_t}} \label{2d-sys-homeostasis-coord_transf_ft2} \\
\end{align}
where $f_{(\cdot)}$ denotes the Fourier transform. We solved for $f_\theta$ which immediately gave us the power spectrum $P_\theta (\omega) \equiv |f_\theta (\omega)|^2$ of $\theta$:
\begin{equation}
P_\theta (\omega) = \frac{\sigma_{NO}^2}{\omega^2 NO_0^2 \tau_{V_t}^2 \lambda^2 + \left( \omega^2 NO_0 \tau_{V_t} + \frac{\gamma \alpha N_{\rm exc.}}{L^2}\right)^2} \label{Pow_Spec_Theta_Final}
\end{equation}
Since we assumed the noise term $\xi$ to be approximately white and gaussian noise with unit variance, its power spectrum is $|f_{\xi}|^2 = 1$.

@@

\subsubsection{Distribution of Firing Rates}\label{Fir_Dist_Section}
Achieving a broad distribution of firing rates among excitatory neurons was the core motivation for the implementation of diffusive homeostasis. Figure \ref{Fir_Rate_Dist_Compare} shows a first result, comparing both homeostatic mechanisms. As expected, non-diffusive homeostasis leads to a sharp distribution of firing rates at 3 Hz. Diffusive homeostasis indeed results in a much broader distribution of mean firing rates. As mentioned in the introduction, a large number of experimental studies have found that distributions of firing rates are not only broadly distributed but well described by a log-normal distribution, which has a non-zero third moment or skewness. By definition, the logarithm of the random variable in question is thus again normally distributed. We plotted the the distribution of decadic logarithms of firing rates in Figure \ref{Fir_Rate_Dist_Compare} (B) to check for this property. In (A), we found a skewness of $\mathrm{v_{\rm diff.} = 0.77}$, in (B) $\mathrm{v_{\rm diff.,log} = -0.49}$. Though this told us that the distribution is ``more symmetric" on a logarithmic scale, it should rather be seen as being neither strictly normally or log-normally distributed.  
\begin{figure}
%%% plot_frequ_dist_mult_sets.py
\includegraphics[width=\textwidth]{../../plots/firing_rate_dists/fir_rate_dist_e_compare.png}
\includegraphics[width=\textwidth]{../../plots/firing_rate_dists/fir_rate_dist_i_compare.png}
\caption{Histograms of mean firing rates over the excitatory/inhibitory population in regular (A/A*) and logarithmic space (B/B*). For diffusive homeostasis ($\mathrm{D=10\, \mu m^2 ms^{-1}}$), the distribution was generated from 10 simulation runs, 1 simulation was used for non-diffusive homeostasis. Mean firing rates were calculated from spikes within $\mathrm{1000-1500\,s}$.}
\label{Fir_Rate_Dist_Compare}
\end{figure}
In contrast, inhibitory cells showed a much clearer distinction: A skewness of $\mathrm{v_{\rm diff.} = 1.52}$ in regular space and $\mathrm{v_{\rm diff.,log} = -0.15}$ in log-space.

@@

A naturally emerging question when altering the diffusion constant is how the firing rate behaves in the absolute limit of infinitely fast diffusion. In fact, this case is quite easy to implement simulation-wise: One simply has to feed all NO-sources into a single scalar variable of NO concentration. This will provide the same NO readout for all excitatory neurons, which means that all excitatory thresholds change at the same rate all the time, only shifting their initial random distribution. Figure \ref{Fir_Rate_Dist_Instant_compare} shows the distribution for this special limiting case. The standard deviation for the excitatory population was $\mathrm{\sigma_{\rm inst.} = 1.79\, Hz}$ and the skewness $\mathrm{v_{\rm inst.} = 1.51}$ ($\mathrm{v_{\rm inst.,log} = -0.52}$), which makes the asymmetry slightly more pronounced than in Figure \ref{Fir_Rate_Dist_Compare}.  
\begin{figure}
%%% frequ_dist_instant_diff.py
\includegraphics[width=\textwidth]{../../plots/firing_rate_dists/instant_diff_fir_dist_e.png}
\includegraphics[width=\textwidth]{../../plots/firing_rate_dists/instant_diff_fir_dist_i.png}
\caption{Distribution of firing rates for instantaneous diffusion for excitatory (A,B) and inhibitory (A*,B*) population. Data was taken from 10 simulations and $\mathrm{1000\,s \leq t \leq 1500\,s}$.}
\label{Fir_Rate_Dist_Instant_compare}
\end{figure}

Summing up the results of this section, we can state that stable network activity can be achieved in the LIF-SORN with diffusive homeostasis while producing a broad distribution of firing rates within the excitatory population. We found that this ensemble of activity is maintained on a single cell level, each cell spiking with approximately poissonian statistics at a constant mean rate. The broadness of distribution could be maintained at a relatively constant level over a wide range of diffusion rates. However, we did not find a distribution as right skewed and heavy tailed as reported by Sweeney et al. and numerous experimental studies.

@@

@@

As an indirect consequence of spatial structure determining activity (as discussed in Section \ref{Fir_Rate_Dist_Discussion_Section}) we found that neurons with a low-density neighborhood have a higher chance of developing these strong outgoing weights. Since this effect is directly linked to the low-density/high-firing-rate relation shown in Figure \ref{Inverse_Dens_vs_Sum_Out_Weights} (C), all previous thoughts regarding plausibility similarly apply to this result: \emph{If} a relationship between density and firing rates existed, we could indeed argue to also expect a relation as shown in Figure \ref{Inverse_Dens_vs_Sum_Out_Weights} (D), since it is known that synapses with above-average firing presynaptic neurons tend to undergo long-term potentiation \cite{Sjoestroem_Syn_Plasticity_2001,Feldman_STDP_2012}.

\newpage
\section{Conclusion and Outlook}\label{Conclusion_Section}
Drawing a conclusion from our results presents itself as a somewhat ambiguous task. On a phenomenological level, our main goal of allowing the LIF-SORN to exhibit a broad, skewed distribution was achieved. However, whether the implications of the diffusive model that revealed itself upon deeper investigations are acceptable in terms of biological plausibility is still left open to debate.

Our analysis of the of the feedback control has shown that the dynamics of the system can be approximately described in a linearized way. As we have speculated in Section \ref{Possible_Modifications_Section}, a better model of homeostatic feedback might eventually require the introduction of nonlinearity, at the cost of a harder to analyze and predict behavior.

From a topological point of view, the implementation of homeostasis can be regarded as a success: Thanks to the broad range of neural activity it allowed for the emergence of highly influential subgroups of excitatory neurons. This---experimentally confirmed---feature was not present in earlier versions of the LIF-SORN. It thus can be regarded as another non-random property of cortical structure that was not hard-coded but naturally emerged from the set of basic rules that constitute dynamics and behavior of our network.

With respect to future use and implementation of diffusive homeostasis one might desire to cover the essential effects in a more simple and abstract formalism. As shown in Section \ref{Mean_Field_Review_Section}, a simple mean field model as presented by Sweeney et al. still carries problems similar to those of single-neuron homeostasis. However, modifications of this model that incorporate our remarks in Section \ref{Possible_Modifications_Section} might be able to recover properties of the full diffusive homeostatic feedback while being more computationally efficient.

\clearpage
\bibliography{test_base}
\bibliographystyle{unsrt}
\end{document}

>>>>

\usepackage[titletoc]{appendix}

@@

\subsubsection{The Role of Firing Rate Heterogeneity---a Dichotomy between Heterogeneity and Control?}\label{Section_Role_Fir_Rate_Het}
Modeling homeostasis necessarily includes a form of error estimation, which requires setting some predefined target that the feedback control will attempt to achieve. Of course this raises the question of how to choose these homeostatic targets. As mentioned in the previous section, former versions of the SORN defined a single global target value. However, experimental studies have found that neuronal populations exhibit a strong heterogeneity within their individual activity. Many statistical analyses found particular evidence for a log-normal like distribution of firing rates \cite{Buzsaki_Fir_Rates_2014,Wohrer_Fir_Rates_2012}. Following these results, one approach of modeling this feature would be to randomly distribute individual targets according to the aforementioned statistics. A problem of this procedure is the absence of any particular deeper physiological motivation or causation. In the light of this conceptual problem, one could, due to a lack of better knowledge, claim that it is reasonable to define a single global target of activity.

@@

Combining these results in a set of equations, we found
\begin{align}
\dot{NO}(t) &= -\lambda NO + \frac{ \left( r_{IP} + \alpha \left( V_{t,\rm pop.}(t)-V_{t,0} \right) \right)  N_{\rm exc.}\frac{{Ca^{2+}_{\rm spike}}^3 \tau_{Ca^{2+}} \mathrm{ln} (2)}{3}}{L^2} + \sigma_{NO} \xi(t) \label{2d-sys-homeostasis1}\\
\dot{V}_{t,\rm pop.}(t) &= \frac{NO-NO_0}{NO_0 \tau_{V_t}}\;. \label{2d-sys-homeostasis2}
\end{align}
Through the coordinate transformations
\begin{align}
n &\equiv NO-NO_0 \label{simplif_coord_transf1}\\
\theta &\equiv V_{t,\rm pop.}(t)-V_{t,0} \label{simplif_coord_transf2}\\
\end{align}
we could simplify the dynamical system:
\begin{align}
\dot{n} &= -\lambda n + \frac{\gamma \alpha N_{\rm exc.}}{L^2}\theta + \sigma_{NO} \xi(t) \label{2d-sys-homeostasis-coord_transf1}\\
\dot{\theta} &= \frac{1}{NO_0 \tau_{V_t}} n \label{2d-sys-homeostasis-coord_transf2}
\end{align}
where we have additionally used
\begin{align}
\gamma &\equiv \frac{{Ca^{2+}_{\rm spike}}^3 \tau_{Ca^{2+}} \mathrm{ln} (2)}{3} \label{2d-sys-fixed-points1}\\
NO_0 &= \frac{nNOS_{\mathrm{ total},0}}{\lambda L^2} \label{2d-sys-fixed-points2}\\
nNOS_{\mathrm{ total},0} &= r_{IP} N_{\rm exc.} \cdot \gamma \; . \label{2d-sys-fixed-points3}
\end{align}
As stated in the beginning, we were interested in finding an analytic expression for the power spectrum of the mean excitatory threshold. We therefore calculated the Fourier transform of \eqref{2d-sys-homeostasis-coord_transf1} and \eqref{2d-sys-homeostasis-coord_transf2}, yielding
\begin{align}
i\omega f_n &= - \lambda f_n + \frac{\gamma \alpha N_{\rm exc.}}{L^2}f_\theta + \sigma_{NO} f_{\xi} \label{2d-sys-homeostasis-coord_transf_ft1}\\
i\omega f_\theta &= \frac{f_n}{NO_0\tau_{V_t}} \label{2d-sys-homeostasis-coord_transf_ft2} \\
\end{align}
where $f_{(\cdot)}$ denotes the Fourier transform. We solved for $f_\theta$ which immediately gave us the power spectrum $P_\theta (\omega) \equiv |f_\theta (\omega)|^2$ of $\theta$:
\begin{equation}
P_\theta (\omega) = \frac{\sigma_{NO}^2}{\omega^2 NO_0^2 \tau_{V_t}^2 \lambda^2 + \left( \omega^2 NO_0 \tau_{V_t} + \frac{\gamma \alpha N_{\rm exc.}}{L^2}\right)^2} \label{Pow_Spec_Theta_Final}
\end{equation}
Appendix \ref{Appendix_Power_Spectrum} contains the full calculation.

@@

\subsubsection{Distribution of Firing Rates}\label{Fir_Dist_Section}
Achieving a broad distribution of firing rates among excitatory neurons was the core motivation for the implementation of diffusive homeostasis. Figure \ref{Fir_Rate_Dist_Compare} shows a first result, comparing both homeostatic mechanisms. As expected, non-diffusive homeostasis leads to a sharp distribution of firing rates at 3 Hz. Diffusive homeostasis indeed results in a much broader distribution of mean firing rates. As mentioned in Section \ref{Section_Role_Fir_Rate_Het}, a large number of experimental studies have found that distributions of firing rates are not only broadly distributed but well described by a log-normal distribution, which has a non-zero third moment or skewness. By definition, the logarithm of the random variable in question is thus again normally distributed. We plotted the the distribution of decadic logarithms of firing rates in Figure \ref{Fir_Rate_Dist_Compare} (B) to check for this property. A Gaussian fit of the histogram in log-space (B) fitted the data reasonably well, with a mean $\mu_{\rm diff.,exc.} = \rm 0.152 \pm  0.002\, log_{10}(Hz)$ and a standard deviation of $\sigma_{\rm diff.,exc.} = \rm 0.152 \pm  0.002\, log_{10}(Hz)$. In (A), we found a skewness of $\mathrm{v_{\rm diff.,exc.,reg.} = 0.77}$, in (B) $\mathrm{v_{\rm diff.,exc.,log.} = -0.49}$. Though this told us that the distribution is ``more symmetric" on a logarithmic scale, the fact that we still find a skewness on the same order as in regular space indicated that it should rather be seen as not perfectly log-normal distributed.  
\begin{figure}
%%% plot_frequ_dist_mult_sets.py
\includegraphics[width=\textwidth]{../../plots/firing_rate_dists/fir_rate_dist_e_compare.png}
\includegraphics[width=\textwidth]{../../plots/firing_rate_dists/fir_rate_dist_i_compare.png}
\caption{Histograms of mean firing rates over the excitatory/inhibitory population in regular (A/A*) and logarithmic space (B/B*). For diffusive homeostasis ($\mathrm{D=10\, \mu m^2 ms^{-1}}$), the distribution was generated from 10 simulation runs, 1 simulation was used for non-diffusive homeostasis. Green curves are non-linear least-squares Gaussian fit onto the distribution resulting from diffusive homeostasis. Mean firing rates were calculated from spikes within $\mathrm{1000-1500\,s}$.}
\label{Fir_Rate_Dist_Compare}
\end{figure}
In contrast, inhibitory cells showed a much clearer distinction: A skewness of $\mathrm{v_{\rm diff.,inh.,reg.} = 1.52}$ in regular space and $\mathrm{v_{\rm diff.,inh.,log.} = -0.15}$ in log-space. Furthermore, the broadness of firing rates was more pronounced, with a Gaussian fit resulting in $\mu_{\rm diff.,inh.} = \rm 0.793 \pm  0.006\, log_{10}(Hz)$ and $\sigma_{\rm diff.,inh.} = \rm 0.190 \pm  0.006\, log_{10}(Hz)$.

@@

A naturally emerging question when altering the diffusion constant is how the firing rate behaves in the absolute limit of infinitely fast diffusion. In fact, this case is quite easy to implement simulation-wise: One simply has to feed all NO-sources into a single scalar variable of NO concentration. This will provide the same NO readout for all excitatory neurons, which means that all excitatory thresholds change at the same rate all the time, only shifting their initial random distribution. Figure \ref{Fir_Rate_Dist_Instant_compare} shows the distribution for this special limiting case. The standard deviation for the excitatory population was $\mathrm{\sigma_{\rm inst.,reg.} = 1.79\, Hz}$ ($\mathrm{\sigma_{\rm inst.,log.} = 0.28\, log_{10}(Hz)}$) and the skewness $\mathrm{v_{\rm inst.,reg.} = 1.51}$ ($\mathrm{v_{\rm inst.,log} = -0.52}$), which makes the asymmetry slightly more pronounced than in Figure \ref{Fir_Rate_Dist_Compare}.  
\begin{figure}
%%% frequ_dist_instant_diff.py
\includegraphics[width=\textwidth]{../../plots/firing_rate_dists/instant_diff_fir_dist_e.png}
\includegraphics[width=\textwidth]{../../plots/firing_rate_dists/instant_diff_fir_dist_i.png}
\caption{Distribution of firing rates for instantaneous diffusion for excitatory (A,B) and inhibitory (A*,B*) population. Red curves are least-squares Gaussian fits ($\mu_{\rm exc.} = \rm 0.420 \pm  0.006\, log_{10}(Hz)$, $\sigma_{\rm exc.} = \rm 0.254 \pm  0.006\, log_{10}(Hz)$, $\mu_{\rm inh.} = \rm 0.761 \pm  0.007\, log_{10}(Hz)$, $\sigma_{\rm inh.} = \rm 0.233 \pm  0.007\, log_{10}(Hz)$). Data was taken from 10 simulations and $\mathrm{1000\,s \leq t \leq 1500\,s}$.}
\label{Fir_Rate_Dist_Instant_compare}
\end{figure}

Summing up the results of this section, we can state that stable network activity can be achieved in the LIF-SORN with diffusive homeostasis while producing a broad distribution of firing rates within the excitatory population. We found that this ensemble of activity is maintained on a single cell level, each cell spiking with approximately poissonian statistics at a constant mean rate. The broadness of distribution could be maintained at a relatively constant level over a wide range of diffusion rates. However, we did not find a distribution as right skewed and heavy tailed as reported by Sweeney et al. Though we will argue in Section \ref{Matrix_Diff_Model_Section} that the quenched random structure of the neurons' positions plays an important role in shaping the distribution of firing rates as well, a potential explanation might be the fact that Sweeney et al. did not impose postsynaptic normalization onto their (static) recurrent excitatory connections, which were randomly generated. A balanced state was achieved by a global scaling factor between excitatory and inhibitory conductances. Therefore, some neurons could potentially be driven by significantly more (or less) synaptic input currents than on average, allowing for greater heterogeneity of firing rates.

@@

Experimental studies have not only considered the presence of neurons with strong influence onto other neurons \cite{Eckmann_Leader_Neurons_2008}, but also found evidence that highly active neurons form subnetworks with an increased connectivity, in the sense that these cells were more likely than average to connect to each other, see \cite{Yassin_Subnetworks_2010}. We tested this in our network by evaluating the connection fraction $\rm CF_{diff.,leader}$ among the top 10\% of excitatory neurons with respect to their average firing rate within $\rm 1000\,s < t < 1500\,s$ and comparing it to the rest of the excitatory population and the same calculation for the non-diffusive case. This returned $\rm CF_{diff.,leader} = 0.1371 \pm 0.0061$, $\rm CF_{diff.,non-leader} = 0.0982 \pm 0.0013$, $\rm CF_{non-diff.,leader} = 0.0954 \pm 0.0044$ and $\rm CF_{non-diff.,non-leader} = 0.0992 \pm 0.0002$. Standard errors were calculated from 5 simulation runs for each homeostatic mechanism, and connectivity matrices were taken from $\rm t = 1500\,s$. This showed that connectivity is indeed increased among the fastest-firing neurons 10\% in the case of diffusive homeostasis. Effenberger et al. reported a similar effect, though differences of connectivity were even more pronounced \cite{Effenberger_2015}.

@@

Though theoretical and experimental studies have reported similarly about the presence of exceptionally ``ìnfluential" neurons \cite{Effenberger_2015,Yassin_Subnetworks_2010,Eckmann_Leader_Neurons_2008}, a lack of a clear concept of ``leader neurons" makes comparisons difficult and should be made with care. On the one hand, one could refer to the finding, that some neurons are more effective in terms of local spike initiation or bursting \cite{Eckmann_Leader_Neurons_2008}. The presence of stronger outgoing weights for some excitatory neurons and an increment of pre-post spiketime correlations within the excitatory population generally support this finding. On the other hand, another aspect one could conceptually associate to ``driver neurons" is the presence of subnetworks of highly active neurons with stronger connectivity \cite{Yassin_Subnetworks_2010}. Though we did find evidence that the connection fraction of a subpopulation of fast-spiking neurons was increased, the meaning and implication of this effect on a functional level needs more investigation. As mentioned in Section \ref{Section_Role_Fir_Rate_Het}, it has been argued that the combination of a large group of slowly firing neurons, which subject to stronger neuronal rewiring, and a small group of fast-firing neurons, forming more stable connections, is required to allow the brain to efficiently process and transmit information while being able to undergo longer lasting learning processes, which require plastic changes in the network structure \cite{Dragoi_2003,Buzsaki_2004,Buzsaki_Fir_Rates_2014}. Thus, one might consider testing the performance of learning tasks that had previously been investigated using the SORN and LIF-SORN under diffusive homeostasis. Potentially, this would allow the network to use the number of available nodes in a more efficient way.

As an indirect consequence of spatial structure determining activity (as discussed in Section \ref{Fir_Rate_Dist_Discussion_Section}) we found that neurons with a low-density neighborhood have a higher chance of developing strong outgoing weights. Since this effect is directly linked to the low-density/high-firing-rate relation shown in Figure \ref{Inverse_Dens_vs_Sum_Out_Weights} (C), all previous thoughts regarding plausibility similarly apply to this result: \emph{If} a relationship between density and firing rates existed, we could indeed argue to also expect a relation as shown in Figure \ref{Inverse_Dens_vs_Sum_Out_Weights} (D), since it is known that synapses with above-average firing presynaptic neurons tend to undergo long-term potentiation \cite{Sjoestroem_Syn_Plasticity_2001,Feldman_STDP_2012}.

\newpage
\section{Conclusion and Outlook}\label{Conclusion_Section}
Our main goal of implementing diffusive homeostasis was to allow a broad distribution of firing rates across the excitatory population of the LIF-SORN, since experimental research has shown this to be a extensively observable phenomenon in the cortex. The implementation of diffusive homeostasis can be regarded as a success with respect to this feature. Furthermore, we did not observe any breakdown of previously studied desirable features of network topology, emerging from the plasticity mechanisms included.

The fact that unwanted oscillations appeared on the first attempts of running the network with diffusive homeostasis forced us to better understand the dynamics of the feedback control and allowed us to contextualize the results into other theoretical research on the stability of neural homeostasis.

Our theoretical considerations regarding the steady state of firing activity allowed us to further clarify the actual determinants of the resulting firing rate statistics. It turned out that, depending on the choice of diffusion parameters, a weighted mixture between the distribution of neural inputs and the spatial configuration determines the resulting statistics of firing rates. The former aspect is in line with common theoretical approaches of explaining firing rate distributions. The latter aspect can be subsumed by the hypothesis that local fluctuations of neural density can influence firing rates and shape their overall distribution. This was a rather unconventional idea, and we had to leave the question of biological plausibility unanswered.

The implementation of diffusive homeostasis led to an interesting observation regarding network structure: Due to the broad range of neural activity it allowed for the emergence of highly influential subgroups of excitatory neurons. This feature was not present in earlier versions of the LIF-SORN. It can be regarded as another non-random property of cortical structure that was not hard-coded but naturally emerged from the set of basic rules that constitute dynamics and behavior of our network. As stated earlier, future research should investigate whether this effect also influences the performance of the SORN in learning tasks.

With respect to future use and implementation of diffusive homeostasis one might also desire to cover the essential effects in a more simple and abstract formalism. As shown in Section \ref{Mean_Field_Review_Section}, a simple mean field model as presented by Sweeney et al. still carries problems similar to those of single-neuron homeostasis. However, modifications of this model that incorporate our remarks in Section \ref{Possible_Modifications_Section} might be able to recover properties of the full diffusive homeostatic feedback while being more computationally efficient.

\bibliography{test_base}
\bibliographystyle{unsrt}

\begin{appendices}

\section{Full Derivation of the Power Spectrum given in Equation \eqref{Pow_Spec_Theta_Final}}\label{Appendix_Power_Spectrum}
Starting with the dynamic equations
\begin{align}
\dot{n} &= -\lambda n + \frac{\gamma \alpha N_{\rm exc.}}{L^2}\theta + \sigma_{NO} \xi(t) \label{Appendix_Equation_1}\\
\dot{\theta} &= \frac{1}{NO_0 \tau_{V_t}} n \label{Appendix_Equation_2}
\end{align}
taking the Fourier transform yields
\begin{align}
i\omega f_n &= - \lambda f_n + \frac{\gamma \alpha N_{\rm exc.}}{L^2}f_\theta + \sigma_{NO} f_{\xi} \label{Appendix_Equation_3} \\
i\omega f_\theta &= \frac{f_n}{NO_0\tau_{V_t}}\;. \label{Appendix_Equation_4}\\
\end{align}
We solve Equation \eqref{Appendix_Equation_4} for $f_n$:
\begin{equation}
f_n = i\omega NO_0 \tau_{V_t} f_\theta \label{Appendix_Equation_5}
\end{equation}
Substituting into Equation \eqref{Appendix_Equation_3} gives
\begin{equation}
-\omega^2 NO_0 \tau_{V_t} f_\theta = - i\omega NO_0 \tau_{V_t} \lambda f_\theta + \frac{\gamma \alpha N_{\rm exc.}}{L^2}f_\theta + \sigma_{NO} f_{\xi} \; . \label{Appendix_Equation_6}
\end{equation}
Solving for $f_\theta$ results in
\begin{equation}
f_\theta = \frac{\sigma_{NO} f_{\xi}}{i\omega NO_0 \tau_{V_t} \lambda - \left(\omega^2 NO_0 \tau_{V_t} + \frac{\gamma \alpha N_{\rm exc.}}{L^2} \right)} \; . \label{Appendix_Equation_7}
\end{equation}
The squared absolute value of this expression is the power spectrum $P_\theta$:
\begin{equation}
P_\theta (\omega) = |f_\theta|^2 = \frac{\sigma_{NO}^2}{\omega^2 NO_0^2 \tau_{V_t}^2 \lambda^2 + \left( \omega^2 NO_0 \tau_{V_t} + \frac{\gamma \alpha N_{\rm exc.}}{L^2}\right)^2} \label{Appendix_Equation_8}
\end{equation}
Note that since we assumed the noise term $\xi$ to be approximately white and Gaussian noise with unit variance, its power spectrum is $|f_{\xi}|^2 = 1$

\end{appendices}

\end{document}
